---
title: "Wisconsin Gerrymandering"
sidebar: false
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(tidyverse)
library(sf)
library(fec16)
library(leaflet)
library(mdsr)
library(mapproj)
library(maps)
library(ggspatial)
library(prettymapr)
library(readr)
```

```{r, include = FALSE}
results_house |>
  group_by(state, district_id) |>
  summarize(N = n()) |>
  print(n=Inf)
```

```{r, include = FALSE}
district_elections <- results_house |>
  mutate(district = parse_number(district_id)) |>
  group_by(state, district) |>
  summarize(
    N = n(), 
    total_votes = sum(general_votes, na.rm = TRUE),
    d_votes = sum(ifelse(party == "DEM", general_votes, 0), na.rm = TRUE),
    r_votes = sum(ifelse(party == "REP", general_votes, 0), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    other_votes = total_votes - d_votes - r_votes,
    r_prop = r_votes / total_votes,  
    winner = ifelse(r_votes > d_votes, "Republican", "Democrat")
  ) |>
  filter(state == "WI") |>
  select(-state)
```

```{r, include = FALSE}
src <- "http://cdmaps.polisci.ucla.edu/shp/districts113.zip"
lcl_zip <- fs::path(tempdir(), "districts113.zip")
download.file(src, destfile = lcl_zip)
lcl_districts <- fs::path(tempdir(), "districts113")
unzip(lcl_zip, exdir = lcl_districts)
dsn_districts <- fs::path(lcl_districts, "districtShapes")

districts <- st_read(dsn_districts, layer = "districts113") |>
  mutate(DISTRICT = parse_number(as.character(DISTRICT)))

wi_shp <- districts |>
  filter(STATENAME == "Wisconsin") |>
  st_transform(4326) |>
  inner_join(district_elections, by = c("DISTRICT" = "district"))
```

```{r, include = FALSE}
ggplot(data = wi_shp, aes(fill = winner)) +
  annotation_map_tile(zoom = 6, type = "osm", progress = "none") + 
  geom_sf(alpha = 0.5) +
  scale_fill_manual("Winner", values = c("Democrat" = "blue", "Republican" = "red")) + 
  geom_sf_label(aes(label = DISTRICT), fill = "white") +
  theme_void()
```

```{r, echo = FALSE}
pal <- colorNumeric(palette = "RdBu", domain = c(0, 1))

leaflet(wi_shp) |>
  addTiles() |>
  addPolygons(
    weight = 1, fillOpacity = 0.8, 
    color = ~pal(1 - r_prop),
    popup = ~paste("District", DISTRICT, "</br>", "Proportion Republican Votes", round(r_prop, 4))
  ) |>
  setView(lng = -90, lat = 45, zoom = 6) |>
  addControl(html = '<h2 style="text-align: center;">2016 WI Congressional Districts</h2>', position = "topright")
```

### WI Gerrymandering

This map provides a detailed choropleth representation of Wisconsin's congressional districts as they were defined in 2016, illustrating the election outcomes by color-coding districts according to the majority political party: blue for Democrats and red for Republicans. The intensity of each color reflects the proportion of votes received by the winning party, with darker hues indicating a larger majority. Noteworthy are the unique shapes of several districts, particularly Districts 3 and 4, which have raised discussions on gerrymandering due to their unusual boundaries and political outcomes. District 3 extends into areas like Stevens Point and Eau Claire, while District 4 encompasses a compact area around Milwaukee. The disparity in district sizes and shapes, such as the expansive District 7, hints at the complex dynamics of political representation and the distribution of liberal and conservative votes across the state. This visualization aids in understanding the geographical and political landscape of Wisconsin, highlighting both the results of the 2016 congressional elections and the ongoing conversations around districting practices.
